DRAFT: v4 \\ Vibration Characterisation of Strollers and\\ Cargo Bicycles for Transporting Infants\\ \vspace{2mm} \large{Including Recommendations for Users, Designers, Manufacturers, and Researchers} \\ \vspace{2mm} \large{Prepared for VeiligheidNL}

### 2025-02-07 main.tex
Updated paper to version 4.

\usepackage{amsmath}  % extra math features
\usepackage{booktabs}  % nice tables
\usepackage[margin=25mm]{geometry}
\usepackage{graphicx} % required for inserting images
\usepackage{hyperref} % clickable hyperlinks
\usepackage{multirow}  % pandas to_latex() uses this
\usepackage{siunitx}  % use for all units
\usepackage{subcaption}  % for subfigures

@@

\title{
  Vibration Characterisation of Strollers and\\
  Cargo Bicycles for Transporting Infants\\
  \vspace{2mm}
  \large{Including Recommendations for Users, Designers, Manufacturers, and Researchers} \\
  \vspace{2mm}
  \large{Prepared for VeiligheidNL}
}
% Gabriele as first, Jason as last, alphabetical for intermediate (discuss in meeting if other method desired
\author{
Gabriele Dell'Orto \and
Brecht Daams \and
Riender Happee \and
Georgios Papaioannou \and
Arjo J. Loeve \and
Jesper Meijerink \and
Thomas Valk \and
Jason K. Moore
}

@@

\color{red}
\section*{\centering Notice}
%
This report or its findings may not be shared outside of VeiligheidNL and
TU~Delft without further review of the methods and findings.
We also need to carefully consider possible adverse effects of dissemination. In
particular we need to consider:
%
\begin{itemize}
  \item System manufacturers being negatively affected by publication.
    Organisations like Consumentenbond typically give manufacturers access to
    draft test results, and give an opportunity to react. We can do this as
    well. We may now remove Brand and type information from this report. In a
    later stage we may test multiple products and then publish a comparison.
  \item Stimulating car use to transport infants and children. We don't want
    newspaper/online headlines which are overly alarming, and later turn out to
    be debatable.
\end{itemize}
\color{black}

\begin{abstract}
  Infants are transported in strollers and in baby seats mounted in cargo
  bicycles. The infants experience whole-body vibration during the associated
  walking and cycling trips, but there is little existing knowledge about
  associated comfort and health effects. To improve this, we measure the seat
  pan acceleration of five strollers and two styles of cargo bicycles with dummy
  infants representing ages 0 to 9 months over six different road surfaces of
  varying road roughness at typical travel speeds. The five strollers included
  three modern systems and two vintage systems with some level of suspension.
  One cargo bicycle had electric support enabling speeds up to 25~\si{\kph}. We
  report average seat pan acceleration for 78 different scenarios and
  investigate the effects of road surface, vehicle, seat type, and travel speed.
  Using the whole-body vibration ISO standards, we show that rough road
  surfaces, amplified by travel speed, may have health risks for travel
  durations as low as \SIrange{10}{30}{\min}. Stroller designs from 50 years ago
  significantly reduce vibrations, relative to most modern strollers we tested,
  indicating the need for return of suspension elements. Cargo bicycles ridden
  at the maximum electric bicycle speed over paver bricks can cause
  accelerations that should likely be avoided except for only the shortest of
  durations until more direct evidence of risks to infants is studied. Standards
  are needed for infant transport testing and evaluation, designers need to
  incorporate better suspension for infants, and research is needed to develop
  clear and direct connections to health and comfort effects.
\end{abstract}

@@

% Don't edit, working to change the following paragraph
In the context of infants transportation with stroller seats, Kok
Siong~\cite{lim2018study} carried out an indoor and an outdoor experiment to
capture the vibrations transmitted onto the seat and backrest of a baby
stroller, using weights from \SIrange{4}{14.14}{\kilo\gram}, and  a child
subject of 10.30~\si{\kilo\gram} as well as a dummy weighing
10.30~\si{\kilo\gram}. 
The comfort levels for the indoor testing were labelled as `Fairly
uncomfortable' to `A little uncomfortable', but the outdoor testing resulted in
`Extremely uncomfortable' and `Fairly uncomfortable' values. They used both a
child and dummy for the experiments, identifying similar patterns. However, the
results still consider ISO~2631-1 standard, which is designed for adults and
specific postures. In a similar research, Sierzputowski~
et~al.~\cite{sierzputowski2021pilot} investigated vibrations induced in a modern
and
an older stroller, when moving on different surfaces: asphalt road, concrete
paving blocks, concrete plates, dirt road, lawn and damaged concrete. The
obtained weighted RMS acceleration values based on ISO~2631-1 were several times
in the range of the highest discomfort (>2~\si{\mps\squared}). 
Another research which included also children as test subjects was conducted by
Okajima~et~al.~\cite{okajima2020dynamic} measuring vibrations that seven
children received, sitting upright in a stroller, when riding over a protrusion
(13x60 mm) with either one wheel or two wheels. 
The children were 3 to 6 years old, boys and girls, weighing
\SIrange{15.94}{19.96}{\kilo\gram} and measuring
\SIrange{98.0}{112.4}{\centi\meter} in length. 
Based on the results, the infant’s head and chest exhibit strong vibrations at
1~\si{\hertz} along all three axes (x,y and z), vibrate easily along the
y-direction at 2~\si{\hertz}, and are less likely to vibrate at more than
8~\si{\hertz}. 
This result is critical as this range of frequencies is very common in bicycles
and strollers. However, they did not assess the comfort and the health risk when
the children were exposed to these vibrations. 
Despite significant research on the effect of whole-body vibrations on
children's comfort, there is limited research on infants.

@@

\subsection{Postures}
%
All tested systems provided full support of back and head and allowed usage in
lying or at least reclined postures. Nine month old infants would be able to sit
erect but will often rest their heads or be asleep. Although sitting upright is
the best posture for children who can sit upright, a more reclined posture was
tested because this allowed measurement of vibration at the head-seat interface
while the head was always supported by the headrest. We tested all systems with
horizontal or reclined postures. The inclination angle with respect to the
ground is listed in Table~\ref{tab:baby-posture} for each condition tested.
%
\begin{table}
  \centering
  \caption{Inclination angles of the IMUs ``Seat Head'' and ``Seat Pan'', per
  each tested configuration. We report the angles defined by the axis of the
  human body reference system (a reference system connected to the dummy, whose
  z-axis is aligned with the body spine and x-axis points to the front of the
  body) and the ground plane. An angle 90~\si{\degree} indicates supine posture,
  -90~\si{\degree} indicates recumbent posture, and 0~\si{\degree} means erect
  posture.}
\begin{tabular}{lllrrr}
\toprule
 &  &  & & \multicolumn{2}{c}{Inclination angle [deg]}  \\
Vehicle Type & Model & Dummy & Baby seat & Seat Head & Seat Pan\\
\midrule
\multirow[t]{6}{*}{Cargo Bicycle} & \multirow[t]{3}{*}{Urban Arrow} & 0 mo & Maxi-Cosi X Joolz & 49
& -43 \\
 &  & 3 mo & Maxi-Cosi X Joolz & 49 & -43 \\
 &  & 3 mo & Melia & 26 & -28 \\
\cline{2-6}
 & \multirow[t]{3}{*}{Keiler} & 0 mo & Maxi-Cosi X Joolz & 49 & -43 \\
 &  & 3 mo & Maxi-Cosi X Joolz & 49 & -43 \\
 &  & 3 mo & Melia & 26 & -28 \\
\cline{1-6} \cline{2-6}
\multirow[t]{5}{*}{Stroller} & Maxi-Cosi Street Plus & 0 mo & Baby cot & 90 & 90\\
\cline{2-6}
 & Maxi-Cosi Street Plus & 9 mo & Baby seat & 50 & -40 \\
\cline{2-6}
 & Babyzen YOYO 0+ & 0 mo & Baby cot & 78 & 78 \\
\cline{2-6}
& Babyzen YOYO 0+ & 9 mo & Baby seat & 45 & -41 \\
\cline{2-6}
& Bugaboo Fox 5 & 0 mo & Baby cot & 90 & 90 \\
\cline{2-6}
 & Bugaboo Fox 5 & 9 mo & Baby seat & 41 & -24\\
\bottomrule
\end{tabular}
  \label{tab:baby-posture}
\end{table}

@@

\subsection{Recommendations}
%
Our conclusions result primarily from the recommended application of ISO~2631-1.
It is of utmost importance to recognize that this standard cautions extending
the use of the standard to situations of which its supporting data was not
derived. The standard is based primarily on pre-1997 studies of adult whole-body
vibration in occupational durations (\SIrange{4}{8}{\hour} daily dose) or during
public transportation.
%
\begin{quote}
  \textbf{The recommendations in standard ISO~2631-1 have not been based on or validated for: 1)
  short durations that dominate stroller and cargo bicycle travel, 2) infants,
  children, or young adults, 3) nor to recumbent or reclined seating postures with the head supported.
  So our recommendations must be taken with utmost caution, at least until more
  research is done to improve the standard guidance.}
\end{quote}
%
Nevertheless, the following list provides recommendations for users,
researchers, designers, and manufacturers based on the findings of this paper
that we believe stand in spite of the standard's limitations:
%
\begin{enumerate}
    \item Stroller, cargo bicycle, and seat manufacturers should test for
    vibration for all expected surface types and ranges of relevant body sizes
    and posture to ensure their designs isolate vibrations well for all use
    conditions. Testing across relevant body sizes is especially important for
    strollers, due to the high mass of the infant relative to the stroller mass.
    \item Manufacturers should collaborate in testing and report their results
    publicly, similar to the automobile industry safety ratings. For the
    combined use of cargo bicycles and baby seats, manufacturers should
    collaborate in this effort.
    \item Manufacturers and scientists should collaborate to develop metrics and
    testing procedures for the long-term goal of a new standard.
    \item Designers should ensure that adequate vibration isolation occurs for
    vehicles that have multiple configurations (e.g. recumbent vs. erect
    seating).
    \item Designers and manufacturers should incorporate better suspension
    systems, as currently occurring vibrations can be drastically reduced.
    Useful information may be derived from past designs with better suspension.
    \item Cycling roads should have a tarmac-like smoothness. Cobblestones
    should be avoided.
    \item Sidewalks should have a tarmac-like smoothness. Cobblestones should be
    avoided for strollers.
    \item Users should limit their speed when walking with strollers or riding
    over surfaces rougher than tarmac, as bumpier surfaces can quickly multiply
    the experienced accelerations.
    \item Users are suggested to limit the duration of transport over surfaces
    rougher than tarmac to periods that do not exceed 10-30 continuous minutes
    depending on the system (existence of suspension, larger tyres or wheels)
    and other countermeasures like adequate non-vibration breaks in between.
    When transporting over tarmac, continuous durations should not exceed
    4~hours and need to include regular breaks, if only to give the infant
    opportunity to move.
    \item Preferably ride any cargo bicycle at a speed of maximal 12~\si{\kph}
    over surfaces rougher than tarmac. E-cargo bicycles can easily reach the
    maximum speed of 25~\si{\kph}. Infants should only be transported for short
    durations when riding over non-tarmac surfaces at these speeds.
    \item Standard ISO~2631-1 is not based on or validated to characterize
    health and comfort effects of vibration for infants or children or for short
    durations or non-erect seating. However, this is the only available standard
    in the literature assessing health and comfort levels on vibrations, and
    could be used as benchmark.
\end{enumerate}

>>>>

\usepackage{amsmath}  % extra math features
\usepackage{booktabs}  % nice tables
\usepackage{draftwatermark}  % use [nostamp] to disable
\usepackage{fancyhdr}
\usepackage[margin=25mm]{geometry}
\usepackage{graphicx} % required for inserting images
\usepackage{hyperref} % clickable hyperlinks
\usepackage{multirow}  % pandas to_latex() uses this
\usepackage{siunitx}  % use for all units
\usepackage{subcaption}  % for subfigures

% fancyhdr settings
\pagestyle{fancy}
\fancyhead[L]{Dell'Orto et. al, Vibration Characterisation of Strollers and Cargo Bicycles for Transporting Infants}
\fancyhead[R]{\thepage}

% watermark settings
\SetWatermarkColor[gray]{0.92}
\SetWatermarkText{Non-disclosure}
\SetWatermarkScale{5}

@@

\title{
  DRAFT: v4 \\
  Vibration Characterisation of Strollers and\\
  Cargo Bicycles for Transporting Infants\\
  \vspace{2mm}
  \large{Including Recommendations for Users, Designers, Manufacturers, and Researchers} \\
  \vspace{2mm}
  \large{Prepared for VeiligheidNL}
}
% Gabriele as first, Jason as last, alphabetical for intermediate (discuss in meeting if other method desired
\author{
Gabriele Dell'Orto \and
Brecht Daams \and
Riender Happee \and
Georgios Papaioannou \and
Arjo J. Loeve \and
Jesper Meijerink \and
Thomas Valk \and
Jason K. Moore~\footnote{corresponding author: j.k.moore@tudelft.nl, +31 (0)15 278 3556}
}

@@

\color{red}
\section*{\centering Notice}
%
This draft report is shared by TU Delft and VeiligheidNL with a number of
societal, industrial and, scientific parties. We aim to submit this to a journal
at the end of February 2025. Your feedback will be appreciated.
\\ \\
This report is shared under NON-DISCLOSURE. If you receive this report you may
share and discuss it with colleagues but do not distribute it further without
our permission. Do not place this document anywhere on internet and do not cite
this version.

\color{black}
\begin{abstract}
  Infants are transported in strollers and in baby seats mounted in cargo
  bicycles. The infants experience whole-body vibration during the associated
  walking and cycling trips, but there is little existing knowledge about
  associated comfort and health effects. To improve this, we measure the seat
  pan acceleration of five strollers and two styles of cargo bicycles with dummy
  infants representing ages 0 to 9 months over six different road surfaces of
  varying road roughness at typical travel speeds. The five strollers included
  three modern systems and two vintage systems with some level of suspension.
  One cargo bicycle had electric support enabling speeds up to 25~\si{\kph}. We
  report average seat pan acceleration for 78 different scenarios and
  investigate the effects of road surface, vehicle, seat type, and travel speed.
  Using the whole-body vibration ISO standards, we show that rough road
  surfaces, amplified by travel speed, may have health risks for travel
  durations as low as \SIrange{10}{30}{\min}. Stroller designs from 50 years ago
  significantly reduce vibrations, relative to most modern strollers we tested,
  indicating the need for return of suspension elements. Cargo bicycles ridden
  at the maximum electric bicycle speed over paver bricks can cause
  accelerations that should likely be avoided except for only the shortest of
  durations until more direct evidence of risks to infants is studied. Standards
  are needed for infant transport testing and evaluation, designers need to
  incorporate better suspension for infants, and research is needed to develop
  clear and direct connections to health and comfort effects.
\end{abstract}

@@

% Don't edit, working to change the following paragraph
In the context of infants transportation with stroller seats, Kok
Siong~\cite{lim2018study} carried out an indoor and an outdoor experiment to
capture the vibrations transmitted onto the seat and backrest of a baby
stroller, using weights from \SIrange{4}{14.14}{\kilo\gram}, and  a child
subject of 10.30~\si{\kilo\gram} as well as a dummy weighing
10.30~\si{\kilo\gram}. 
The comfort levels for the indoor testing were labelled as `Fairly
uncomfortable' to `A little uncomfortable', but the outdoor testing resulted in
`Extremely uncomfortable' and `Fairly uncomfortable' values. They used both a
child and dummy for the experiments, identifying similar patterns. However, the
results still consider ISO~2631-1 standard, which is designed for adults and
specific postures. In a similar research, Sierzputowski~
et~al.~\cite{sierzputowski2021pilot} investigated vibrations induced in a modern
and
an older stroller, when moving on different surfaces: asphalt road, concrete
paving blocks, concrete plates, dirt road, lawn and damaged concrete. The
obtained weighted RMS acceleration values based on ISO~2631-1 were several times
in the range of the highest discomfort ($>$2~\si{\mps\squared}). 
Another research which included also children as test subjects was conducted by
Okajima~et~al.~\cite{okajima2020dynamic} measuring vibrations that seven
children received, sitting upright in a stroller, when riding over a protrusion
(13x60 mm) with either one wheel or two wheels. 
The children were 3 to 6 years old, boys and girls, weighing
\SIrange{15.94}{19.96}{\kilo\gram} and measuring
\SIrange{98.0}{112.4}{\centi\meter} in length. 
Based on the results, the infant’s head and chest exhibit strong vibrations at
1~\si{\hertz} along all three axes (x,y and z), vibrate easily along the
y-direction at 2~\si{\hertz}, and are less likely to vibrate at more than
8~\si{\hertz}. 
This result is critical as this range of frequencies is very common in bicycles
and strollers. However, they did not assess the comfort and the health risk when
the children were exposed to these vibrations. 
Despite significant research on the effect of whole-body vibrations on
children's comfort, there is limited research on infants.

@@

\subsection{Postures}
%
All tested systems provided full support of back and head and allowed usage in
lying or at least reclined postures. Nine month old infants would be able to sit
erect but will often rest their heads or be asleep. Although sitting upright is
the best posture for children who can sit upright, a more reclined posture was
tested because this allowed measurement of vibration at the head-seat interface
while the head was always supported by the headrest. We tested all systems with
horizontal or reclined postures. The inclination angle with respect to the
ground is listed in Table~\ref{tab:baby-posture} for each condition tested.
%
\begin{table}
  \centering
  \caption{Inclination angles of the IMUs ``Seat Head'' and ``Seat Pan'', per
  each tested configuration. We report the angles defined by the axis of the
  human body reference system (a reference system connected to the dummy, whose
  z-axis is aligned with the body spine and x-axis points to the front of the
  body) and the ground plane. An angle 90~\si{\degree} indicates supine posture (lying horizontally),
  -90~\si{\degree} indicates recumbent posture, and 0~\si{\degree} means erect
  posture.}
\begin{tabular}{lllrrr}
\toprule
 &  &  & & \multicolumn{2}{c}{Inclination angle [deg]}  \\
Vehicle Type & Model & Dummy & Baby seat & Seat Head & Seat Pan\\
\midrule
\multirow[t]{6}{*}{Cargo Bicycle} & \multirow[t]{3}{*}{Urban Arrow} & 0 mo & Maxi-Cosi X Joolz & 49
& -43 \\
 &  & 3 mo & Maxi-Cosi X Joolz & 49 & -43 \\
 &  & 3 mo & Melia & 26 & -28 \\
\cline{2-6}
 & \multirow[t]{3}{*}{Keiler} & 0 mo & Maxi-Cosi X Joolz & 49 & -43 \\
 &  & 3 mo & Maxi-Cosi X Joolz & 49 & -43 \\
 &  & 3 mo & Melia & 26 & -28 \\
\cline{1-6} \cline{2-6}
\multirow[t]{5}{*}{Stroller} & Maxi-Cosi Street Plus & 0 mo & Baby cot & 90 & 90\\
\cline{2-6}
 & Maxi-Cosi Street Plus & 9 mo & Baby seat & 50 & -40 \\
\cline{2-6}
 & Babyzen YOYO 0+ & 0 mo & Baby cot & 78 & 78 \\
\cline{2-6}
& Babyzen YOYO 0+ & 9 mo & Baby seat & 45 & -41 \\
\cline{2-6}
& Bugaboo Fox 5 & 0 mo & Baby cot & 90 & 90 \\
\cline{2-6}
 & Bugaboo Fox 5 & 9 mo & Baby seat & 41 & -24\\
\bottomrule
\end{tabular}
  \label{tab:baby-posture}
\end{table}

@@

\subsection{Recommendations}
Our conclusions result primarily from the recommended application of ISO~2631-1.
It is of utmost importance to recognize that this standard cautions extending
the use of the standard to situations of which its supporting data was not
derived. The standard is based primarily on pre-1997 studies of adult whole-body
vibration in occupational durations (\SIrange{4}{8}{\hour} daily dose) or during
public transportation.
\begin{quote}
  \textbf{The recommendations in standard ISO~2631-1 have not been based on or
  validated for infants, children, or young adults nor for recumbent or reclined
  seating postures with the head supported. Additionally, the admissible
  durations for health are not validated for daily short durations or shorter
  lifetime accumulation that dominate stroller and cargo bicycle travel with
  infants. So our recommendations must be taken with caution, at least until
  more research is done to improve the standard guidance.}
\end{quote}
Nevertheless, the following list provides recommendations for users,
researchers, designers, and manufacturers based on the findings of this paper
that we believe stand in spite of the standard's limitations:
\begin{enumerate}
    \item Stroller, cargo bicycle, and seat manufacturers should test for
    vibration for all expected surface types and ranges of relevant body sizes
    and posture to ensure their designs isolate vibrations well for all use
    conditions. Testing across relevant body sizes is especially important for
    strollers, due to the high mass of the infant relative to the stroller mass.
    \item Manufacturers should collaborate in testing and report their results
    publicly, similar to the automobile industry safety ratings. For the
    combined use of cargo bicycles and baby seats, manufacturers should
    collaborate in this effort.
    \item Manufacturers and scientists should collaborate to develop metrics and
    testing procedures for the long-term goal of a new standard.
    \item Designers should ensure that adequate vibration isolation occurs for
    vehicles that have multiple configurations (e.g. recumbent vs. erect
    seating).
    \item Designers and manufacturers should incorporate better suspension
    systems, as currently occurring vibrations can be drastically reduced.
    Useful information may be derived from past designs with better suspension.
    \item Cycling roads should have a tarmac-like smoothness. Cobblestones
    should be avoided.
    \item Sidewalks should have a tarmac-like smoothness. Cobblestones should be
    avoided for strollers.
    \item Users should limit their speed when walking with strollers or riding
    over surfaces rougher than tarmac, as bumpier surfaces can quickly multiply
    the experienced accelerations.
    \item Users are suggested to limit the duration of transport over surfaces
    rougher than tarmac to periods that do not exceed 10-30 continuous minutes
    depending on the system (existence of suspension, larger tyres or wheels)
    and other countermeasures like adequate non-vibration breaks in between.
    When transporting over tarmac, continuous durations should not exceed
    4~hours and need to include regular breaks, if only to give the infant
    opportunity to move.
    \item Preferably ride any cargo bicycle at a speed of maximal 12~\si{\kph}
    over surfaces rougher than tarmac. E-cargo bicycles can easily reach the
    maximum speed of 25~\si{\kph}. Infants should only be transported for short
    durations when riding over non-tarmac surfaces at these speeds.
    \item Standard ISO~2631-1 is not validated to characterize health and
    comfort effects of vibration for infants or children, for short travel
    durations, for vibration accumulation over only a small number of years, or
    for non-erect seating. However, this is the only available standard in the
    literature assessing health and comfort levels on vibrations, and should be
    used as benchmark.
\end{enumerate}
